The Strong Gravitational Lens Finding Challenge

### 2018-12-04 response.tex

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\author{R.B. Metcalf}
\begin{document}
Referee Report

{\bf 
Overall, this work is an important milestone towards exploring the usefulness of strong gravitational lenses in many applications in astrophysics, and after significant copy-editing and clarifications should be published.  I particularly agree that the emphasis on attempting to understand the most successful lens-candidate identification approaches and on the role of false positives is important.

The submitted manuscript is full of incorrectly assigned or compiled references.  The Sections’ organizational structure is incomplete and mis-referenced. Furthermore, the paper seems not to have been proof-read carefully as there are many minor grammatical and typesetting errors that are avoidable and distracting.  This includes typos within equations.  All of these must be corrected and scrupulously proof-read for re-submission.

The long-term goal of this collaboration of maturing algorithms for delivering automated gravitational lens detection across datasets as diverse as Euclid, LSST, and SKA, is timely and an important effort for the community.  Developing such flexible infrastructure is going to be challenging, and as the authors themselves point out, the realism of the simulated datasets will be an essential element. To that end, it makes sense that the present work gives a taste of what this may look like, by considering two parallel efforts in tandem, focusing on the single-band space-based Euclid and the multiple-band ground-based Kilo-Degree Survey (KiDS) datasets. 
}

{\bf
The paper seems to have been written towards an audience that is already closely familiar with gravitational lensing topics, which misses an opportunity to have a broader impact. For example, in stating “These take the form of multiple images of a quasar and Einstein rings surrounding a galaxy and of highly distorted arcs seen through galaxy groups and clusters.” conflates several lens geometries and situations in a way that e.g. a student would need to disentangle. 
}

The paper is generally targeted toward astronomers, but the point is taken.  The first paragraph of the introduction has been expanded to better describe gravitational lensing.  We do not wish to take up to much space describing the whole field when this can be read about elsewhere.

{\bf
In the abstract (but nowhere else it seems) there is a summary result that “several methods are able to identify upwards of half the lenses”.  It seems that this is a statement of significance or interpretation and so it seems incomplete.  Or is it simply a statement of fact? 
}

This is referred to in paragraph 4 of section 6.2.  It is a conclusion drawn from figures 9 and 10.  We are not sure what the referee is asking here.

{\bf
What was the rationale for the 20.000 size of the training sets and the choice of the 101x101 pixels? 
}

No particularly rational reason.  The 101x101 pixels originally was the size of some early stage cutouts from KIDS.  20,000 was a convenient number for processing time and storage reasons.  We have added a comment on this.

{\bf
The challenge dataset is stated as having 100.000 objects, and in one place only there is a reference that in “the case of the multi-band ground-based set this was 400,000 images.”  There’s never another reference to the 400.000, however, so this setup description is confusing or incomplete. 
}

It is clearly stated several times that there were 4 bands in the ground based set which makes 400,000 images for 100,000 objects.

{\bf
Understanding how the simulation set was prepared is significant for understanding the entire project and its results.  The Millennium simulation was used to identify galaxies.  The generally available Millennium catalog is at redshift zero, rather than at the various lens-plane redshifts given.  If some epoch-matching was made, please explain this, or justify the use of redshift-zero galaxy properties.  I do note a description of the adopted catalog being projected to a series of twenty lens planes with the total deflection calculated through a sum, but remain unclear about the process.  I also do not follow what “A NFW profile is fit to the [total mass, size, and half mass radius] parameters” actually means for this application.  As galaxies identified within Millennium were used, these are generally sub-critical to strong lensing, unless an additional central baryonic steepening has been applied. Please clarify.  
}

We did not use just the zero redshift galaxies.  The details of the light-cone construction is given in Overzier, et al. 2013 MNRAS 428, 778.  They used a semi-analytic model to populate the halos with galaxies.  The galaxies are accounted for in both light (section 3 paragraph 8 through 11, formerly 7 through 10) and mass 
(section 3 paragraphs 12 and 13, formerly 11-12).  We have tried to make this clearer  by changing the first paragraph of the section.

{\bf
In the Introduction’s description of imaging searches for lenses, references to past work is quite incomplete, and it would be appreciated to have these developed further.  For example, while there is a passing reference to the Marshall et al HAGGLeS work, this is arguably a significant precursor to the present paper so it is surprising that it is not discussed at all.  I think that giving depth of context to readers is critical for this paper. 
}

We have added a little more discussion of this paper in the introduction and conclusion.

{\bf
The dismissal of spectroscopic surveys for lenses seems flippant in addition to being incomplete. There are several spectroscopic lens projects after SLACS, and if the spectroscopic approach does return “well defined and pure” samples, it seems that at least some reference to the potential of DESI and PFS is expected to be in comparison to imaging surveys. 
}

We added a reference to these instruments in the introduction.

{\bf
It is good to see arc finders included in this challenge, though there is scant quantitative information about how the analysis was set up with these techniques to be able to judge or comment on them. In a comment related to the CNNs’ below, what are development or advancement areas that the Arc Finder community already knew, or learned based on this work, that will be addressed in the future?  
}

?????

{\bf
The Machine Learning / Support Vector Machine approaches seem clear enough, though these sections also include spare or undefined information.  I happen to know that AUC means area under a curve and how it pertains to the ROC exercise at hand, but this is quite mysterious and incomplete as presented. This can and must be improved significantly in the re-editing of the manuscript.}

Yes, unfortunately the ROC is not defined until section 6.1.  We have put in a reference to this section so that the reader knows where to find the definition if they don't know it.

{\bf Incidentally, the “AUC of 0.88 for the space set and 0.95 for the ground set” stated in section 5.3.1 are different numbers from what are given in Tables 3 \& 4, further casting doubt on the validation of all results here.  
}

That is the AUC on the training set not the challenge set.

{\bf
The Convolutional Neural Networks general description is well done.  This is a complex and rich topic that is certainly still developing within astrophysics, and this section gives enough sense of the uncertainty and nuance behind exactly how the network methodologies may be set up for particular analyses.  Note that “neurones” are “neurons”, suggesting that additional care is needed in verifying the presentation in this section.   
}

"neurones" is the British spelling.  The spelling has been standardized to "neurons" and "neuron".

{\bf
How the training sets are handled by each CNN approach is very interesting reading, and there is a great deal to digest about the choices for convolution layers, neuron numbers, and classification metrics.  The variety of how rotation/reflection/spatial transformations of the training sets are treated is interesting, as is the varied use of multiple bands’ images or not.  Given the focus on CNN methods, the reader may expect some greater discussion after the lens-finding conclusions on what areas of advancement or improvement emerged from this exercise, even at high level.   This is offered as an observation.  
}

?????

{\bf
In fact, some of the algorithm descriptions are rather spare in their details, and so difficult to judge or intercompare.  The reader (and this referee) is forced to take much at face value, which perhaps is acceptable for the scope of an “overview” report such as this.  It is a little disappointing, all the same.  A uniform template in the structure of each contributor’s report could have helped mitigate that, so that similar context or informational questions would at least have a chance of being reported.  The additional descriptive detail given by some of the contributors, in particular LASTRO, GAMOCLASS, and CAST are appreciated. 
}

??????

{\bf
The receiver operating characteristic (ROC) curve technique is referred to via its acronym in many places in the manuscript without explanation or definition, and even in the Results Section the description is 
}

Added this.

{\bf
There is a brief discussion about rare “abnormal” objects including merging and irregular galaxies potentially being “the greatest challenge” for a full realistic application. 
An accurate representation of these objects is promised for future work. Can the authors elaborate how this may be done? 
}

We are working on this.  There are several approaches we are considering.  They would 
involve randomizing existing images.

{\bf
This referee agrees forcefully with the statement in the Conclusions \& Discussion section that application of SVM and CNN methods have “some danger of over-fitting to the training set”, and that there is enormous weight on how realistic the simulations are.  This discussion draws no conclusions, however, about how confident the community may reasonable expect to become with this or future work, and how to get there.  A brief discussion of this should be included.  
}

This is a good question and one I don't think was fully answered by the challenge.  You might have 
thought that visual inspection is the gold standard, but it seems that you can do better.  But maybe 
that is just over-fitting.  It might be that unsupervised learning techniques have a part to play.

A little discussion has been added to the conclusion, but any more would be pure speculation.

{\bf
The overall conclusion or claim that ground-based multi-band data are more successful than low SNR single-band space-based data is a quite strong and even disturbing statement.  The way it is cast here, it seems to imply that we should not bother with space-based (Euclid) data at all for detecting strong lenses.  Is that the ultimate message the authors wish to convey to their audience?  Are there factors of solid angle, depth, or data stability and robustness that may be factors in the near future if not just at present?  The final statement on the importance of colour information for Euclid, whether from the infrared bands that were not presently included or by the combination with other ground-based data, is well done. Is future work planned that will include these dimensions? 
}

It is true that mono-band space based data might not be a big step up for this purpose, but 
multi-band data exists and even if it is of lower resolution, as it will be for Euclid, it should 
make space based searches more successful than ground based ones.  Yes, we are fallowing 
up with Euclid UV simulations.  We have tried to change this impression by adding a note to the abstract.

\end{document}

### 2018-12-04 paper.tex
Edit.

\abstract{
%\begin{abstract}
Large scale imaging surveys will increase the number of galaxy-scale strong lensing candidates by maybe three orders of magnitudes beyond the number known today.  Finding these rare objects will require picking them out of at least tens of millions of images and deriving scientific results from them will require quantifying the efficiency and bias of any search method.  To achieve these objectives automated methods must be developed.
Because gravitational lenses are rare, objects reducing false positives will be particularly important.  We present a description and results of an open gravitational lens finding challenge.  Participants were asked to classify 100,000 candidate objects as to whether they were gravitational lenses or not with the goal of developing better automated methods for finding lenses in large data sets.   A variety of methods were used including visual inspection, arc and ring finders, support vector machines (SVM) and convolutional neural networks (CNN).  We find that many of the methods will be easily fast enough to analyse the anticipated data flow.  In test data, several methods are able to identify upwards of half the lenses after applying some thresholds on the lens characteristics such as lensed image brightness, size or contrast with the lens galaxy without making a single false-positive identification.  This is significantly better than direct inspection by humans was able to do.  Having multi-band, ground based data is found to be better for this purpose than single-band space based data with lower noise and higher resolution, suggesting that multi colours data is crucial additional.
The most difficult challenge for a lens finder is differentiating between rare irregular and ring-like face-on galaxies and true gravitational lenses.  The degree to which the efficiency and biases of lens finders can be quantified largely depends on the realism of the simulated data on which the finders are trained.
}%\end{abstract}

@@

Strong gravitational lenses are rare cases where a distant galaxy or quasar is aligned so closely with a foreground galaxy or cluster of galaxies  that the gravitational field of the foreground object creates multiple, highly distorted images of the background object.  These take the form of multiple images of a quasar and Einstein rings surrounding a galaxy and of highly distorted arcs seen through galaxy groups and clusters.  The first strong lens was discovered in 1979 by \citet{1979Natur.279..381W} and since then many valuable scientific uses have been found for them.  They have been used to study how dark matter is distributed in galaxies and clusters \citep[e.g.][]{1991ApJ...373..354K,2001ApJ...554.1216C,2002ApJ...568L...5K,2003ApJ...587..143R,2003ApJ...583..606K,2005MNRAS.360.1333W,2005ApJ...623...31D,2009MNRAS.392..945V,2016MNRAS.463.3115T} and  to measure the Hubble constant and other cosmological parameters 
\citep[e.g.][]{1964MNRAS.128..307R,1992ARAandA..30..311B,2000ApJ...544...98W,2013ApJ...766...70S,2016AandARv..24...11T}.  
Their magnification has been used as a natural telescope to observe otherwise undetectable objects at high redshift \citep[e.g][]{2007ApJ...671.1196M,2017MNRAS.464.4823B,2016ApJ...833..264S}.   They have put limits on the self interaction of dark matter and on the theory of gravitation \citep{0004-637X-606-2-819}.  Through microlensing they have been used to study the structure of quasars \citep{2008ApJ...689..755M,2008ApJ...673...34P,2011ApJ...729...34B}.   To access this wealth of information the first step is to find the lenses.

@@

Spectroscopic searches for galaxy scale lenses have also been done by looking for high redshift stellar lines in the spectra of lower redshift large galaxies.  Notably this was done in the Sloan Lens ACS (SLACS) survey, producing a relatively well defined and pure sample of Einstein ring lenses \citep{2006MNRAS.369.1521W,2006ApJ...638..703B,2012ApJ...744...41B,2015MNRAS.449.3441S}.  However, spectroscopy is telescope-time consuming and for the foreseeable future we are not likely to have spectroscopic surveys that cover anywhere near the number of objects as the planned imaging surveys.

Some automated algorithms have been developed in the past to detect lenses in images.  These have been designed to detect arc-like 
features \citep{2006astro.ph..6757A,2007AandA...472..341S, 2017AandA...597A.135B} and rings 
\citep{2014ApJ...785..144G,2014AandA...566A..63J}.  They have been applied to survey data and found of order 200 lenses \citep{2007AandA...461..813C,2012ApJ...749...38M,2016AandA...592A..75P}.  Another approach has been to use a lens modeling code to fit a model to all the candidates and pick out the ones that fit the model well \citep{2009ApJ...694..924M,2017arXiv170401585S}.

@@

In each case, a training set of 20,000 images in each band was provided for download at any time along with a key giving some properties of the object including whether it was a gravitational lens.  Each image was $101\times101$ pixels.  The participants were free to download these sets and train their algorithms on them.  To enter the contest, the participants needed to register with a team name at which point they would be given a unique key and the address of a test data set.  These data sets contained 100,000 candidates.  In the case of the multi-band ground-based set this was 400,000 images. The participants had 48 hours to upload a classification of all candidates consisting of a score between 0 and 1 0 signifying the lowest confidence that it is a lens and 1 signifying the highest confidence that it is a lens.  This ranking could have been a simple binary (0 or 1) classification  or it could have been a continuous range of numbers representing the probability of being a lens or it could have been a finite number of confidence levels.

@@

Creating the mock images started with a cosmological N-body simulation, in this case it was the Millennium simulation \citep{2009MNRAS.398.1150B}.   A catalog of dark matter halos within a light-cone was constructed within the Millennium Observatory project \citep{2013MNRAS.428..778O}.   The challenge sets were based on a 1.6 sq.deg. light cone extending out to redshift $z=6$.  The halos are found with a friends-of-friends algorithm and characterised by a total mass, size and half mass radius.  They include subhalos of larger halos.  The halo catalog is read into the GLAMER lensing code \citep{2014MNRAS.445.1942M,2014MNRAS.445.1954P} to do all the ray-tracing.    Within this code a Navarro, Frenk \& White (NFW) \citep{1996ApJ...462..563N} profile is fit to the three parameters given above.  The halos are projected onto a series of 20 lens planes  and the deflection angle at any point on each plane are calculated by summing the effects of all the halos with a hybrid tree method.  In this way the halos have the mass, concentration and clustering properties from the N-body simulation, but within each strong lens the mass resolution is not limited by the original simulation, but follows the analytic mass profile.   An additional mass component that will be discussed later is added to each halo to represent the stellar mass.

@@

Our final feature extraction procedure first applied a polar transform to each image in order to exploit the edge detection of the Gabor filter, picking out tangential components typical of galaxy-galaxy lensing. Each image was then convolved with several Gabor filters of varying frequency and rotation (see Fig.~\ref{gaborring}). Stability selection methods were used to investigate the classification performance using different combinations of filters. The responses for each Gabor filter when applied to each image were measured by calculating statistical moments for each filtered image. These moments formed  our final input data on which the SVM could be trained and applied. We used brute-force optimisation methods to select a non-linear SVM containing radial basis function (RBF) kernel and tuned a small set of regularisation hyperparameters to achieve good generalisation performance. During training and testing, our final and best scores achieved when testing on the lens finding challenge training data were an AUC of 0.88 for the space set and 0.95 for the ground set. Classification was performed using a modest desktop PC.

@@

In deriving a good figure of merit for evaluating lens finding algorithms one needs to take into account the particular nature of this problem.
The traditional method for evaluating a classification algorithm is with the receiver operating characteristic curve, or  ROC (see section \ref{sec:figure_of_merit}) curve.  This is a plot of the true positive rate (TPR) and the false positive rate (FPR).  In this case these are defined as
\begin{align}
{\rm TPR} &= \frac{\textrm{ number of true lenses classified as lenses}}{\textrm{ total number of true lenses}} \\
{\rm FPR} &= \frac{\textrm{number of non-lenses classified as lenses}}{\textrm{ total number of non-lenses}}
\end{align}
The classifier generally gives a probability of a case being a lens, $p$, in which case a threshold is set and everything with $p$ greater is classified as a lens and everything smaller is classified as not a lens.  The TPR and FPR are then plotted as a curve parametrised by this threshold.  At $p=1$ all of the cases are classified as non-lenses and so TPR=FPR=1 and at $p=0$ all of the cases are classified as lenses so TPR=FPR=0.  These points are always added to the ROC curve.  If the classifier made random guesses then the ratio of lenses to non-lenses would be the same as the ratio of the number of cases classified as lens to the number of cases classified as non-lenses and so TPR=FPR.  The better a classifier is the smaller the FPR and the larger the TPR so the further away from this diagonal line it will be.  When a classifier provides only a binary classification or a discrete ranking, the ROC connects the endpoints to the discrete points found by using each rank as a threshold.

@@

We have done a fairly good job of determining that lenses can be identified in a population of fairly "normal" galaxies.  It is the rare "abnormal" objects that pose the greatest challenge.  When real KiDS data was used in the simulations the classifiers were all less accurate and it was only human inspection that found the one jackpot lens (a double Einstein ring with two background sources) in the data.  Things like ring galaxies, tidal tails in merging galaxies and irregular galaxies can be mistaken for lenses and were not well represented in the simulated data.  Accurately reproducing these objects will be an objective of future work.

It was surprising to some of the authors how well CNN and SVM methods did relative to human inspection.  These methods find differences in the classes of images that are not obvious to a human and can classify things as lenses with high confidence where a human would have doubt. 
This ability comes with some danger of over fitting to the training set however.  The distinguishing characteristics might only be a property of simulated data and not of real data.  In principle, SVM methods might potentially mitigate this somewhat because with them one can choose which features to use based on knowledge of the properties of irregular galaxies or ring galaxies for example.  This has yet to be shown however.
The confidence one will have in these machine learning methods is really based on the confidence one has in the realism of the simulations.

>>>>

\abstract{
%\begin{abstract}
Large scale imaging surveys will increase the number of galaxy-scale strong lensing candidates by maybe three orders of magnitudes beyond the number known today.  Finding these rare objects will require picking them out of at least tens of millions of images and deriving scientific results from them will require quantifying the efficiency and bias of any search method.  To achieve these objectives automated methods must be developed.
Because gravitational lenses are rare, objects reducing false positives will be particularly important.  We present a description and results of an open gravitational lens finding challenge.  Participants were asked to classify 100,000 candidate objects as to whether they were gravitational lenses or not with the goal of developing better automated methods for finding lenses in large data sets.   A variety of methods were used including visual inspection, arc and ring finders, support vector machines (SVM) and convolutional neural networks (CNN).  We find that many of the methods will be easily fast enough to analyse the anticipated data flow.  In test data, several methods are able to identify upwards of half the lenses after applying some thresholds on the lens characteristics such as lensed image brightness, size or contrast with the lens galaxy without making a single false-positive identification.  This is significantly better than direct inspection by humans was able to do.  Having multi-band, ground based data is found to be better for this purpose than single-band space based data with lower noise and higher resolution, suggesting that multi colour data is crucial.  Multi-band space based data will be superior to ground based data.
The most difficult challenge for a lens finder is differentiating between rare irregular and ring-like face-on galaxies and true gravitational lenses.  The degree to which the efficiency and biases of lens finders can be quantified largely depends on the realism of the simulated data on which the finders are trained.
}%\end{abstract}

@@

Strong gravitational lenses are rare cases where a distant galaxy or quasar is aligned so closely with a foreground galaxy or cluster of galaxies  that the gravitational field of the foreground object creates multiple, highly distorted images of the background object.   The first strong lens was discovered in 1979 by \citet{1979Natur.279..381W} and since then several hundred of them have been found.
When the lens is an individual galaxy and the source a quasar, there are two, four or five distinct images of the source.  The time-delay between images,  the magnification ratios between images and the image positions can all be used to model the mass distribution of the lens and measure cosmological parameters.  When the lens is a cluster of galaxies the images of background galaxies are multiplied and distorted into many thin arcs.   When the lens is an individual galaxy and the background source also a galaxy, the lensed images can take the form of a partial or complete ring seen around or through the lens galaxy, an {\it Einstein ring}.

Strong lenses have to date provided very valuable scientific information.  They have been used to study how dark matter is distributed in galaxies and clusters \citep[e.g.][]{1991ApJ...373..354K,2001ApJ...554.1216C,2002ApJ...568L...5K,2003ApJ...587..143R,2003ApJ...583..606K,2005MNRAS.360.1333W,2005ApJ...623...31D,2009MNRAS.392..945V,2016MNRAS.463.3115T} and  to measure the Hubble constant and other cosmological parameters 
\citep[e.g.][]{1964MNRAS.128..307R,1992ARAandA..30..311B,2000ApJ...544...98W,2013ApJ...766...70S,2016AandARv..24...11T}.  
Their magnification has been used as a natural telescope to observe otherwise undetectable objects at high redshift \citep[e.g][]{2007ApJ...671.1196M,2017MNRAS.464.4823B,2016ApJ...833..264S}.   They have put limits on the self interaction of dark matter and on the theory of gravitation \citep{0004-637X-606-2-819}.  Through microlensing they have been used to study the structure of quasars \citep{2008ApJ...689..755M,2008ApJ...673...34P,2011ApJ...729...34B}.   To expand on this wealth of information we must study more lenses.  The first step in doing this is to find more of these rare objects.

@@

Spectroscopic searches for galaxy scale lenses have also been done by looking for high redshift stellar lines in the spectra of lower redshift large galaxies.  Notably this was done in the Sloan Lens ACS (SLACS) survey, producing a relatively well defined and pure sample of Einstein ring lenses \citep{2006MNRAS.369.1521W,2006ApJ...638..703B,2012ApJ...744...41B,2015MNRAS.449.3441S}.  
New spectrographs such as the Dark Energy Spectroscopic Instrument  (DESI) \citep{2016arXiv161100036D} and Subaru Prime Focus Spectrograph (PFS) \citep{2016SPIE.9908E..1MT} have the potential to expanding spectroscopic lens searches greatly.
However, spectroscopy is telescope-time consuming and for the foreseeable future we are not likely to have spectroscopic surveys that cover anywhere near the number of objects as the planned imaging surveys.

Some automated algorithms have been developed in the past to detect lenses in images.  These have been designed to detect arc-like 
features \citep{2006astro.ph..6757A,2007AandA...472..341S, 2017AandA...597A.135B} and rings 
\citep{2014ApJ...785..144G,2014AandA...566A..63J}.  They have been applied to survey data and found of order 200 lenses \citep{2007AandA...461..813C,2012ApJ...749...38M,2016AandA...592A..75P}.

\cite{2009ApJ...694..924M} pioneered an automated technique for finding Einstein ring type lenses that uses a lens modeling code to fit a model to all candidates and picks out the ones that fit the model well \citep[see also][]{2017arXiv170401585S}.   This approach has the attractive feature that it distinguishes lenses from non-lenses by their similarity to what we expect a lens to look like and priors can be put on the model parameters that are physically motivated, unlike the next category of finders below.  Challenges arise in making the modeling fast enough and automatic enough to handle large 
data sets while allowing it to be flexible enough to find unusual lens configurations.  The YattaLens entrant to this challenge was of this type.

@@

In each case, a training set of 20,000 images in each band was provided for download at any time along with a key giving some properties of the object including whether it was a gravitational lens.  Each image was $101\times101$ pixels.  These specifications were not of particular significance except that the image size would encompass almost all galaxy-galaxy lenses and that the number of images (including ones with and without noise, lens and source) was not too large.
The participants were free to download these sets and train their algorithms on them.  To enter the contest, the participants needed to register with a team name at which point they would be given a unique key and the address of a test data set.  These data sets contained 100,000 candidates.  In the case of the multi-band ground-based set this was 400,000 images. The participants had 48 hours to upload a classification of all candidates consisting of a score between 0 and 1 0 signifying the lowest confidence that it is a lens and 1 signifying the highest confidence that it is a lens.  This ranking could have been a simple binary (0 or 1) classification  or it could have been a continuous range of numbers representing the probability of being a lens or it could have been a finite number of confidence levels.

@@

Creating the mock images started with a cosmological N-body simulation, in this case the Millennium simulation \citep{2009MNRAS.398.1150B}.   A catalog of dark matter halos and galaxies  within a light-cone was constructed within the Millennium Observatory project \citep{2013MNRAS.428..778O}.   The challenge sets were based on a 1.6 sq.deg. light cone extending out to redshift $z=6$ using all the simulation snapshots.  The halos are found with a friends-of-friends algorithm and characterised by a total mass, size and half mass radius.  They include subhalos of larger halos.  The halos where populated with galaxies based on their merger history using the semi-analytic model  (SAM) of \cite{2011MNRAS.413..101G}.

The halo catalog were read into the GLAMER lensing code \citep{2014MNRAS.445.1942M,2014MNRAS.445.1954P} to do all the ray-tracing.    Within this code a Navarro, Frenk \& White (NFW) \citep{1996ApJ...462..563N} profile is fit to the three parameters given above to represent the dark matter component of the lens.  The halos are projected onto a series of 20 lens planes  and the deflection angle at any point on each plane are calculated by summing the effects of all the halos with a hybrid tree method.  In this way the halos have the mass, concentration and clustering properties from the N-body simulation, but within each strong lens the mass resolution is not limited by the original simulation, but follows the analytic mass profile.   An additional mass component that will be discussed later is added to each halo to represent the stellar mass.

@@

Our final feature extraction procedure first applied a polar transform to each image in order to exploit the edge detection of the Gabor filter, picking out tangential components typical of galaxy-galaxy lensing. Each image was then convolved with several Gabor filters of varying frequency and rotation (see Fig.~\ref{gaborring}). Stability selection methods were used to investigate the classification performance using different combinations of filters. The responses for each Gabor filter when applied to each image were measured by calculating statistical moments for each filtered image. These moments formed  our final input data on which the SVM could be trained and applied. We used brute-force optimisation methods to select a non-linear SVM containing radial basis function (RBF) kernel and tuned a small set of regularisation hyperparameters to achieve good generalisation performance. During training and testing, our final and best scores achieved when testing on the  training data were an area under the ROC curve of 0.88 for the space set and 0.95 for the ground set. Classification was performed using a modest desktop PC.

@@

In deriving a good figure of merit for evaluating lens finding algorithms one needs to take into account the particular nature of this problem.
The traditional method for evaluating a classification algorithm is with the receiver operating characteristic curve, or  ROC (receiver operating characteristic) curve.  This is a plot of the true positive rate (TPR) and the false positive rate (FPR).  In this case these are defined as
\begin{align}
{\rm TPR} &= \frac{\textrm{ number of true lenses classified as lenses}}{\textrm{ total number of true lenses}} \\
{\rm FPR} &= \frac{\textrm{number of non-lenses classified as lenses}}{\textrm{ total number of non-lenses}}
\end{align}
The classifier generally gives a probability of a case being a lens, $p$, in which case a threshold is set and everything with $p$ greater is classified as a lens and everything smaller is classified as not a lens.  The TPR and FPR are then plotted as a curve parametrised by this threshold.  At $p=1$ all of the cases are classified as non-lenses and so TPR=FPR=1 and at $p=0$ all of the cases are classified as lenses so TPR=FPR=0.  These points are always added to the ROC curve.  If the classifier made random guesses then the ratio of lenses to non-lenses would be the same as the ratio of the number of cases classified as lens to the number of cases classified as non-lenses and so TPR=FPR.  The better a classifier is the smaller the FPR and the larger the TPR so the further away from this diagonal line it will be.  When a classifier provides only a binary classification or a discrete ranking, the ROC connects the endpoints to the discrete points found by using each rank as a threshold.

@@

We have done a fairly good job of determining that lenses can be identified in a population of fairly "normal" galaxies.  It is the rare "abnormal" objects that pose the greatest challenge.  When real KiDS data was used in the simulations the classifiers were all less accurate and it was only human inspection that found the one jackpot lens (a double Einstein ring with two background sources) in the data.  Things like ring galaxies, tidal tails in merging galaxies and irregular galaxies can be mistaken for lenses and were not well represented in the simulated data.  Accurately reproducing these objects will be an objective of future work.  This might be done by including more real images in the challenge or images based on real images with some random elements added.

It was surprising to some of the authors how well CNN and SVM methods did relative to human inspection.  These methods find differences in the classes of images that are not obvious to a human and can classify things as lenses with high confidence where a human would have doubt. 
This ability comes with some danger of over fitting to the training set however.  The distinguishing characteristics might only be a property of simulated data and not of real data.  In principle, SVM methods might potentially mitigate this somewhat because with them one can choose which features to use based on knowledge of the properties of irregular galaxies or ring galaxies for example.  This has yet to be shown however.  Methods based on fitting with a lens modeling code \citep{2009ApJ...694..924M,2017arXiv170401585S} might also help to mitigate this problem.
The confidence one will have in the machine learning methods is really limited by the confidence one has in the realism of the simulations.   It might be useful in the future to have a challenge without a training set.  This might more clearly reveal the presence of over-fitting.  
It would also be useful to include more real images or images more closely based on real images.
